EBASE manuscript

### 2023-01-23 manu-draft.qmd
minor text edits

The underlying idea behind most gas exchange parameterizations for metabolism estimates is that the transfer of gases of relatively low solubility, like oxygen, is proportional to the concentration gradient in a thin film of water at the air–water interface. It is also assumed that the water at the top of the film is saturated with respect to the atmosphere, so that the concentration gradient across the thin film can be approximated as the difference between oxygen concentration measured in the surface water layer, $C_d$, and the saturation (or equilibrium) concentration $C_{Sat}$, which is usually calculated as a function of water temperature and salinity (the negative sign is needed because positive flux is set to be from water to air):

@@

The metabolism estimates and their parameters returned by the Bayesian routine implemented in JAGS are affected by the prior distributions assigned to each.  As for the BASE model, relatively uninformed prior distributions following a normal Gaussian distribution are used by default, although the priors can be changed based on previous knowledge of parameters specific to an ecosystem or as informed by other metabolic modeling approaches.  The default priors were based on those that returned reasonable parameter estimates using a known time series created with forward-feeding oxygen mass balance inputs to @eq-ebase.  [Fwoxy R package, @Arriola22]. All of the physical parameters were held constant, e.g., wind speed, temperature, and salinity.  The other required inputs were the $a$, $r$, and $b$ parameters, chosen as 0.2 $\auni$, 20 $\runi$, and 0.251 $\buni$, respectively.  An initial oxygen concentration was provided as 250 mmol/m$^3$ and the time series was created by forward addition of the inputs. As such, the dissolved oxygen output as change per unit time was modelled by EBASE using the constant the same physical parameters as input. Reasonable prior distributions by EBASE were identified (Equations [-@eq-aprior], [-@eq-rprior], and [-@eq-bprior]) based on the ability to return the known parameters:

>>>>

A common theory behind most gas exchange parameterizations for metabolism estimates is that the transfer of gases of relatively low solubility, like oxygen, is proportional to the concentration gradient in a thin film of water at the air–water interface. It is also assumed that the water at the top of the film is saturated with respect to the atmosphere, so that the concentration gradient across the thin film can be approximated as the difference between oxygen concentration measured in the surface water layer, $C_d$, and the saturation (or equilibrium) concentration $C_{Sat}$, which is usually calculated as a function of water temperature and salinity (the negative sign is needed because positive flux is set to be from water to air):

@@

The metabolism estimates and their parameters returned by the Bayesian routine implemented in JAGS are affected by the prior distributions assigned to each.  As for the BASE model, relatively uninformed prior distributions following a normal Gaussian distribution are used by default, although the priors can be changed based on previous knowledge of parameters specific to an ecosystem or as informed by other metabolic modeling approaches.  The default priors were based on those that returned reasonable parameter estimates using a known time series created with forward-feeding oxygen mass balance inputs to @eq-ebase.  [Fwoxy R package, @Arriola22]. All of the physical parameters were held constant, i.e., wind speed, temperature, and salinity.  The other required inputs were the $a$, $r$, and $b$ parameters, chosen as 0.2 $\auni$, 20 $\runi$, and 0.251 $\buni$, respectively.  An initial oxygen concentration was provided as 250 mmol/m$^3$ and the time series was created by forward addition of the inputs. As such, the dissolved oxygen output as change per unit time was modelled by EBASE using the constant the same physical parameters as input. Reasonable prior distributions by EBASE were identified (Equations [-@eq-aprior], [-@eq-rprior], and [-@eq-bprior]) based on the ability to return the known parameters:
